### 2025-04-16 Paper/manuscript.tex
add sequence figs

\documentclass[fleqn,10pt,lineno]{manuscript}
%%\usepackage{setspace}
%%\doublespacing
\usepackage{soul}
\usepackage[utf8]{inputenc}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\title{A Structural and Functional Bioinformatics Study of QTY-designed Retinylidene Proteins}

\author[1]{Siqi Pan}
\author[2]{Shuguang Zhang}
\affil[1]{Shanghai World Foreign Language Academy, 400 Baihua Street, Shanghai 200233, China}
\affil[2]{Lab of Molecular Architecture, Media Lab, Massachusetts Institute of Technology, 77 Massachusetts Avenue, Cambridge, MA 02139, USA}

\corrauthor[2]{Shuguang Zhang}{Shuguang@MIT.EDU}

\keywords{Keyword1; Keyword2; Keyword3}
\begin{abstract}
This is the abstract!
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}

Intro

* Families of opsins - vertebrate vs. bacterial

* General features of vertebrate opsin; structure and function

* Activation mechanism of rhodopsin

* Expression, function of each opsin

* General features of bacterial opsin; structure, function, applications

* Why solubilize

* History of solubilizing studies of rhodopsin and bacteriorhodopsin

* Intro to AlphaFold

* Existing QTY studies

* Intro to GROMACS

* Existing rhodopsin bioinformatics studies

\section*{Results and Discussion}

Results

* discuss the QTY code

* describe and explain Table1

* describe and explain Fig1

* describe and explain Fig2
	> I need more discussion here

* describe and explain Fig3

* discuss AlphaFold3 predictions

* describe, explain, discuss MD results (Fig3 and Fig4)
	> I need more discussion here

* future scopes and potential applications

* conclusion

\section*{Methods}

Methods

* protein sequences UniProt

* AlphaFold3 server

* superimposition (PDB, AlphaFold, PyMOL)

* Structure visualization (PyMOL, ChimeraX)

* MD simulation (GROMACS, etc.; detailed params; analysis techniques)

\section*{Supplementary Material}

Can be found at...

\section*{Data Availability Statement}

Can be found at...

\section*{Author contributions}

A.O., A.T. and A.F. conceived the presented idea; A.O. wrote the main manuscript; All authors have read and agreed to the published version of the manuscript.

\section*{Financial Support}

This work was supported ... For the purpose of open access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this submission.

\section*{Acknowledgments}

Thank...

\section*{Competing Interests}

The authors declare no conflict of interest.

\section*{Ethics Statement}

No ethics issues...

\bibliography{references}

\begin{table}[]
	\centering
	\caption{TWO}
	\label{tb:2}
	\begin{tabular}{cccccc}
		\toprule
		Name     & RMSD (\AA) & pI   & Mw (kDa) & TM variation (\%) & Overall variation (\%) \\
		\toprule
		OPSG     &                 & 8.90 & 40.58    &                   &                        \\
		OPSG QTY & 0.468           & 8.82 & 40.85    & 41.86             & 19.78                  \\
		\midrule
		OPSR     &                 & 8.89 & 40.57    &                   &                        \\
		OPSR QTY & 0.611           & 8.83 & 40.76    & 40.12             & 18.96                  \\
		\midrule
		OPSB     &                 & 8.75 & 38.72    &                   &                        \\
		OPSB QTY & 0.486           & 8.66 & 39.30    & 46.51             & 23.19                  \\
		\midrule
		OPSD     &                 & 6.20 & 38.89    &                   &                        \\
		OPSD QTY & 0.559           & 6.20 & 39.29    & 46.58             & 21.55                  \\
		\midrule
		OPN3     &                 & 9.29 & 44.87    &                   &                        \\
		OPN3 QTY & 0.454           & 9.18 & 45.47    & 49.12             & 20.90                  \\
		\midrule
		OPN4     &                 & 9.35 & 52.64    &                   &                        \\
		OPN4 QTY & 0.307           & 9.19 & 53.08    & 50.34             & 15.48                  \\
		\midrule
		OPN5     &                 & 9.11 & 39.73    &                   &                        \\
		OPN5 QTY & 0.555           & 9.02 & 40.04    & 45.58             & 18.93                  \\
		\midrule
		RGR      &                 & 8.34 & 31.87    &                   &                        \\
		RGR QTY  & 0.537           & 8.29 & 32.39    & 42.86             & 21.65                  \\
		\midrule
		OPSX     &                 & 8.77 & 37.42    &                   &                        \\
		OPSX QTY & 0.548           & 8.72 & 37.46    & 40.96             & 20.18                  \\
		\midrule
		BACR     &                 & 4.75 & 26.92    &                   &                        \\
		BACR QTY & 0.448           & 4.75 & 27.40    & 46.67             & 25.30                  \\
		\midrule
		BACH     &                 & 5.34 & 26.96    &                   &                        \\
		BACH QTY & 0.296           & 5.34 & 27.50    & 44.19             & 30.04                  \\
		\bottomrule
	\end{tabular}
\end{table}

\end{document}