### 2025-04-22 Paper/Source/manuscript.tex
Switch to Overleaf TeX. Add RGR and RRH to all OPN superimposition

\documentclass[fleqn,10pt,lineno]{manuscript}
%%\usepackage{setspace}
%%\doublespacing
\usepackage{soul, hyperref}
\usepackage[utf8]{inputenc}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\title{A Structural and Functional Bioinformatics Study of QTY-designed Opsins}

\author[1]{Siqi Pan}
\author[2]{Shuguang Zhang}
\affil[1]{Shanghai World Foreign Language Academy, 400 Baihua Street, Shanghai 200233, China}
\affil[2]{Lab of Molecular Architecture, Media Lab, Massachusetts Institute of Technology, 77 Massachusetts Avenue, Cambridge, MA 02139, USA}

\corrauthor[2]{Shuguang Zhang}{Shuguang@MIT.EDU}

\keywords{Keyword 1; Keyword 2; Keyword 3}
\begin{abstract}

This study investigates the QTY design of 9 human opsins and 3 microbial opsins, all of which bind retinal as the ligand.

\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}

\textbf{* Families of opsins - animal vs. microbial}

\citep{Spudich_2000}

Retinylidene proteins are light-sensing proteins that bind to retinal as a chromophore. 
Prevalence of these proteins - occur in many organisms, from archaea to vertebrates. 
Evolutionarily distinct, convergence; GPCR vs ion pumps or channels.

\textbf{* General features of animal opsin; structure and function; activation mechanism of rhodopsin; may include existing rhodopsin bioinformatics studies?}

Class A GPCR, 7TM, NPxxY motif, outward movement of TM6. \citep{Sakmar_2002, Zhou_2019}
296 Lys in TM7 \citep{Guhmann_2022}.
Lysine-retinal schiff base linkage, conserved evolutionarily. 
Delocalized electrons in retinal, receive photon, high activation rate. 
11-cis to all-trans isomerization.
RHO: dark >> batho >> lumi >> meta I <<>> meta II; proton-transfer pathway. 
TM6 outward. 
Activates G-protein, initiate secondary messenger cascade.

\textbf{* Expression, function of each animal opsin}

The OPN family preferentially binds 11-cis-retinal and catalyzes its isomerization to all-trans-retinal via a retinochrome mechanism. 
Evolutionarily distinct animal opsins \citep{Terakita_2005, Shichida_2009}

OPN1MW (Medium-Wave-sensitive Opsin 1), OPN1LW (Long-Wave-sensitive Opsin 1) and OPN1SW (Short-Wave-sensitive Opsin 1) are proteins expressed in cone photoreceptors in the retina and are responsible for color vision. OPN1MW has maximum absorption at wavelength 530nm (green), OPN1LW at wavelength 560nm (red), and OPN1SW at wavelength 420nm (blue). They are responsible for various color vision defects. Deutanopia, a partial colorblindness characterized by a dichromasy in which red and green are confused, is caused by variants affecting OPN1MW. Protanopia, a partial colorblindness similar to deutanopia, is caused by variants affecting OPN1LW. Tritanopia, a color blindness characterized by selective deficiency of blue spectral sensitivity, is owed to variants of OPN1SW. Blue cone monochromacy, an X-linked congenital cone dysfunction syndrome, is caused by the absence of functional OPN1MW and OPN1LW. Finally, cone dystrophy 5 is an X-linked cone dystrophy characterized by loss of visual acuity, macular lesions, and color vision.

OPN2 (Opsin 2), also known as rhodopsin, is expressed in retinal rod photoreceptors and essential for light transduction and vision at low light intensity. It is correlated with retinitis pigmentosa 4, characterized by retinal pigments deposits, loss of night vision, loss of peripheral visual field, and eventual loss of central visual field. Congenital stationary night blindness, a non-progressive autosomal dominant impairment of night vision, is also attributed to mutation of OPN2. 
RHO: rhodopsin is exemplary; earliest opsin studied, earliest crystal structure of GPCR \citep{Palczeski_2000} bovine rhodopsin.

OPN3 (Opsin 3), also known as encephalopsin or panopsin, is a GPCR activated via ultraviolet A light. 
Regulates melanogenesis in melanocytes. 
Plays a role in melanocyte survival through regulation of intracellular Ca levels. 
Regulates apoptosis via Cyt c release and activation of caspase cascade. 
Keratinocyte differentiation in response to blue light. 
Plays a role in light'mediated glucose uptake, mitochondrial respiration and fatty acid metabolism in brown adipocyte tissues. 
May be involved in photorelaxation of airway smooth muscle cells via blue light.

OPN4 (Opsin 4), also known as melanopsin, is expressed in ipRGC (intrinsically photosensitive Retinal Ganglion Cells) in the ganglion cell layer in the retina. Absorption wavelength??? It is responsible for pupillary reflex, photoentrainment, optokinetic visual tracking response, and other non-image-forming responses to light.

OPN5 (Opsin 5), also known as neuropsin, is a GPCR activated via ultraviolet A light. 
Emission peaks at 380nm (UVA) and 470nm (blue). 
Required for the light-response in the inner plexiform layer, and contributes to the regulation of the light-response in the nerve fiber layer. 
Involved in local corneal and retinal circadian rhythm photoentrainment via modulation of UVA light-induced phase-shift of the retina clock. 
Circadian photoreceptor in the outer ear.

RGR (RPE-retinal GPCR) is expressed in the RPE (retinal pigmented epithelium) and Muller cells in the retina. Unlike the aforementioned OPN family, RGR preferentially binds all-trans-retinal and may catalyze its isomerization via a retinochrome-like mechanism. FUNCTION?

RRH (RPE-derived Rhodopsin Homolog), also known as peropsin, is localized in the microvilli of RPE that surround photoreceptor outer segments. It may play a role in RPE physiology either by detecting light directly or by monitoring the concentration of retinal. FUNCTION?

\textbf{* General features of microbial opsin; structure, function, applications}

7TM? 
Ion channels or pumps? 
Optogenetics. 
Other applications??

\textbf{* Expression, function of each microbial opsin}

\citep{Findlay_1986, Zhang_2011}
BACR, or bacteriorhodopsin, is a 7TM protein that forms a trimer light-driven proton pump.Studied... Applications...
BACH, or halorhodopsin, is a 7TM protein that forms a trimer light-driven chloride pump. Studied... It is activated by yellow light and is used as a tool of inhibition in optogenetics. More applications...
ChR2, or channelrhodopsin 2, is a 7TM protein that forms a dimer light-activated sodium channel. Studied... It is activated by blue light and is used as a tool of excitation in optogenetics. More applications...
All-trans to 13-cis? 
BACH \citep{Baselt_1989, Kouyama_2018}. BACR \citep{Taguchi_2023}. ChR2 \citep{Nagel_2005, Ardevol_2018, Xin_2023}.

\textbf{* History of solubilizing studies of rhodopsin and bacteriorhodopsin}

RESEARCH.

"Recently, researchers built on top of ProteinMPNN to devise SolubleMPNN trained on only soluble proteins, which was applied to engineer soluble variants of bacteriorhodopsin, successfully converting a membrane protein into a soluble one, while maintaining its core function and ligand-binding ability (Nikolaev et al., 2024)."

\textbf{* Why solubilize}

“Recently, we have asked if the QTY code is applicable to other retinylidene proteins. The retinylidene proteins are all integral membrane proteins with seven transmembrane alpha-helices embedded in a lipid bilayer. Therefore, because of the hydrophobic properties of transmembrane domains, they are not water-soluble without the aid of detergents. We wanted to see if the QTY code could be utilized to design water-soluble variants of these retinylidene proteins.”

\textbf{* Existing QTY studies}

"Instead of taking a computational approach, we applied the QTY code to systematically engineer watersoluble analogs with reduced hydrophobicity in membrane proteins. The QTY concept was inspired by high-resolution (1.5\AA) electron density maps, which revealed structural similarities between hydrophobic and polar amino acids leucine (L) vs glutamine (Q); isoleucine (I)/valine (V) vs threonine (T); and phenylalanine (F) vs tyrosine (Y) (Zhang et al., 2018; Zhang \& Egli 2022; Tegler et al., 2020). In our previous experiments, using the simple and straightforward QTY code, we successfully bioengineered detergent-free chemokine (Zhang et al., 2018; Qing et al., 2019; Tegler et al., 2020), cytokine receptors (Hao et al., 2020) and bacterial histidine kinase (Li et al 2024). After these detergent-free membrane proteins were expressed and purified, these QTY analogs demonstrated structural stability, retained their ligand-binding capabilities and intact four enzymatic activities, making them ideal candidates for further studies and use them as antigens to generate therapeutic monoclonal antibodies."

\textbf{* Intro to AlphaFold}

"In May 2024, AlphaFold was upgraded to version 3 as AlphaFold 3, featuring an enhanced diffusion-based architecture that enables accurate prediction of multiple structures of protein complexes. Additionally, AlphaFold 3 extends its capabilities beyond protein structure prediction to include DNA, RNA, and small molecules including ligands and other proteins (Abramson et al., 2024)"

\textbf{* Intro to GROMACS}

RESEARCH.

\textbf{* Overview of this paper}

SUMMARY.

\section*{Results and Discussion}

Results

* discuss the QTY code

* describe and explain Table1

* describe and explain Fig1

* describe and explain Fig2
	- I need more discussion here

* describe and explain Fig3

* discuss AlphaFold3 predictions

* describe, explain, discuss MD results (Fig3 and Fig4)
        - no changes to the NPxxY motif or retinal-binding Lys
	- I need more discussion here

* future scopes and potential applications

* conclusion

\section*{Methods}

Methods

* protein sequences UniProt

* AlphaFold3 server

* superimposition (PDB, AlphaFold, PyMOL)

* Structure visualization (PyMOL, ChimeraX)

* MD simulation (GROMACS, etc.; detailed params; analysis techniques)

\section*{Supplementary Material}

The supplementary material can be found at...

\section*{Data Availability Statement}

The data for ... can be found at...

\section*{Author contributions}

Detailed author contributions

\section*{Financial Support}

No funding was received for this project.

\section*{Acknowledgements}

Thanks to ... for ...

\section*{Competing Interests}

The authors declare no conflict of interest.

\section*{Ethics Statement}

There are no ethics issues related to the research in this paper. No animal or human data...

\bibliography{references}

\begin{table}[htbp]
	\centering
	\caption{Protein characteristics}
	\label{tb:characteristics}
	\includegraphics[width=\linewidth]{Figures/characteristics.jpg}
\end{table}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/sequences.jpg}
	\caption{Protein sequence alignments}
	\label{fig:sequences}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/superimposition-human.jpg}
	\caption{Superimposition of human retinylidene proteins}
	\label{fig:humansup}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/superimposition-microbial.jpg}
	\caption{Superimposition of microbial retinylidene proteins}
	\label{fig:microbialsup}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/pairwise.jpg}
	\caption{Pairwise comparison of human opsins}
	\label{fig:pairwise}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/hydrophobicity.jpg}
	\caption{Surface hydrophobicity}
	\label{fig:hydrophobicity}
\end{figure}

\end{document}